Adaptive Sequential Model Selection

### 2015-07-09 paper.tex
rob

\begin{table}[ht]
\centering
\begin{tabular}{rrrrr}
  \hline\\
 Step& Predictor& Naive pvalue& Saturated pvalue& Max-t pvalue\cr
  \hline\\
1 & 3 & 0.000 & 0.000 & 0.000 \\ 
  2 & 9& 0.000 & 0.000 & 0.000 \\ 
  3 & 4 & 0.000 & \color{red}{0.054} & 0.000 \\ 
  4 & 20& 0.000 & { 0.333} & 0.003 \\ 
  5 & 37& 0.002 & 0.756 & 0.043 \\ 
  6 & 7 & 0.003 & 0.254 & 0.037 \\ 
  7 & 2& 0.000 & 0.003 & 0.000 \\ 
  8 & 19 & 0.020 & 0.028 &  \color{red}{0.343} \\ 
  9 & 11 & 0.112 & 0.548 & 0.937 \\ 
  10 & 52& 0.214 & 0.911 & 0.907 \\ 
  11 & 57& 0.049 & 0.000 & 0.954 \\ 
  12 & 24& 0.222 & 0.000 & 0.996 \\ 
  13 & 5& 0.183 & 0.054 & 0.789 \\ 
  14 & 49 & 0.158 & 0.333 & 1.000 \\ 
  15 & 30 & 0.212 & 0.756 & 1.000 \\ 
  16 & 46 & \color{red}{0.159} & 0.254 & 0.000 \\ 
  17 & 59 & 0.270 & 0.003 & 0.004 \\ 
  18 & 33& 0.116 & 0.028 & 1.000 \\ 
  19 & 62 & 0.081 & 0.548 & 0.668 \\ 
  20 & 27 & 0.170 & 0.911 & 1.000 \\

   \hline\\
\end{tabular}
\caption[tab:diab]{\em Forward stepwise regression for the diabetes data: naive p-values from forward stepwise regression, 
p-values from the ``saturated'' model, and the ``max-t'' p-values that are presented in this paper. The red annotation indicates the stopping step
prescribed by {\tt forwardStop}. For the max-t p-values, this gives exact FDR control at the 10\% level.}
\label{tab:diab}
\end{table}
The table shows the naive p-values from forward stepwise regression, that do not account for the selection, as well as the
p-values from the ``saturated'' model, and the ``max-t'' p-values that are presented in this paper.
The latter two sets of p-values test someone different hypotheses:  the p-values from the saturated model test whether the partial regression of the predictor just added is zero,
what we call the ``incremental null''
In contrast, the max-t pvalues test whether the current model is adequate, that is, all predictors not in the model  have coefficients equal to zero,
the ``complete null''. (The issue of what hypothesis to test is discussed in section \ref{sec:whichnull}.)
Now it turns out that the max-t test if often more powerful that the saturated model test, and importantly, the max-t pvalues are exactly
independent after we reach the point the model sequence when the complete null is true. This independence means that we can apply
sequential FDR exact stopping rules, such as ``forwardStop'' (\citet{gsell2013sequential}). When applied to the max-t pvalues at FDR 0.10, {\tt forwardStop}
suggested stopping at step 8. In constant, the saturated lvalues gave stopping a point of 3, with no guarantee of FDR control
due to the lack of independence of the p-values.

An outline of this paper is as follows:
In Section \ref{sec:genericSetting} we setup the problem and give some important definitions. In Section \ref{sec:sequential} we review recently proposed methods for
testing ordered hypothesis using from a sequence of p-values. These method require that the p-values be uniform and independent under the null, and we derive general conditions
to ensure this property. In Section \ref{sec:singleStep}  we  show how to construct valid selective $p$-values for a single step in the path and derive the max-t test
for  forward stepwise regression and the Next-$\lambda$ test for the lasso. Section \ref{sec:computation}  describes our computational strategy for estimating the p-values using accept-reject and hit and run sampling. We also relate the method to parametric bootstrap sampling. We provide a simulation study in Section \ref{sec:simulation} and briefly discuss applications of our framework to
principal components analysis in section\ref{sec:pac}. The paper ends with a Discussion.

@@

\medskip
\begin{example}[The selective max-Z test]
Consider the forward stepwise path (\ref{eq:forwardDef_start}---\ref{eq:forwardDef_end}) in the case of linear regression, that is 
$\ell(\theta,Y)=||Y-X\theta||^2$.  Assume that the error variance $\sigma^2$ is known.
 Given an active set of variables $E$ in our current model $M$, the partial regression coefficients for predictors not in $E$ are
\begin{equation}
  z_{j,E}(Y) = \frac{(\proj_E^\perp X_j)'Y}
  {\sigma\|\proj_{E}^\perp X_j\|_2}.
\end{equation}
The max-$z$-statistic for active set $E$  is
\begin{equation}\label{eq:maxT}
  U_E(Y) = \max_{j \notin E} \left|t_{j,E}(Y)\right|,
\end{equation}
and we use $U_{E_{k-1}}(Y)$ as our test statistic at step $k$. We compute a p-value for this test
using the conditional law 
\begin{equation}\label{eq:selModel_marg0}
\L\left(X_{j_k}'Y \mid E_{k-1}, \; X_{E_{k-1}}'Y\right),
\end{equation}
If we reject the hypothesis, we add to $E$ the predictor that maximized $U_E(Y)$ and continue.
This is the procedure that used to obtain the p-values in Table \ref{tab:diab}  \RTcomment{ Rob- make sure you fix table 1 to make the correct computation}

>>>>

\begin{table}[ht]
\centering
\begin{tabular}{rrrrr}
  \hline\\
 Step& Predictor& Naive pvalue& Saturated pvalue& Max-t pvalue\cr
  \hline\\
1 & 3& 0.00 & 0.00 & 0.00 \\ 
  2 & 9 & 0.00 & 0.00 & 0.00 \\ 
  3 & 4 & 0.00 & \color{red}{0.05} & 0.00 \\ 
  4 & 20 & 0.00 & 0.33 & 0.01 \\ 
  5 & 37 & 0.00 & 0.76 & 0.05 \\ 
  6 & 7 & 0.00 & 0.25 & 0.05 \\ 
  7 & 2 & 0.00 & 0.00 & 0.00 \\ 
  8 & 19 & 0.02 & 0.03 & \color{red}{0.37} \\ 
  9 & 11 & 0.11 & 0.55 & 0.95 \\ 
  10 & 52 & 0.21 & 0.91 & 0.92 \\ 
  11 & 57 & 0.05 & 0.37 & 0.21 \\ 
  12 & 24 & 0.22 & 0.15 & 0.00 \\ 
  13 & 5 & 0.18 & 0.07 & 0.10 \\ 
  14 & 49 & 0.16 & 0.97 & 0.75 \\ 
  15 & 30& 0.21 & 0.15 & 0.05 \\ 
  16 & 46 & \color{red}{0.16} & 0.05 & 0.53 \\ 
  17 & 59 & 0.27 & 0.45 & 0.68 \\ 
  18 & 33 & 0.12 & 0.71 & 0.80 \\ 
  19 & 62 & 0.08 & 0.40 & 0.48 \\ 
  20 & 27 & 0.17 & 0.60 & 0.50 \\ 
   \hline\\
\end{tabular}
\caption[tab:diab]{\em Forward stepwise regression for the diabetes data: naive p-values from forward stepwise regression, 
p-values from the ``saturated'' model, and the ``max-t'' p-values that are presented in this paper. The red annotation indicates the stopping step
prescribed by {\tt forwardStop}. For the max-t p-values, this gives exact FDR control at the 10\% level.}
\label{tab:diab}
\end{table}
The table shows the naive p-values from forward stepwise regression, that do not account for the selection, as well as the
p-values from the ``saturated'' model, and the ``max-t'' p-values that are presented in this paper.
The latter two sets of p-values test someone different hypotheses:  the p-values from the saturated model test whether the partial regression of the predictor just added is zero,
what we call the ``incremental null''
In contrast, the max-t pvalues test whether the current model is adequate, that is, all predictors not in the model  have coefficients equal to zero,
the ``complete null''. (The issue of what hypothesis to test is discussed in section \ref{sec:whichnull}.)
Now it turns out that the max-t test if often more powerful that the saturated model test, and importantly, the max-t pvalues are exactly
independent after we reach the point the model sequence when the complete null is true. This independence means that we can apply
sequential FDR exact stopping rules, such as ``forwardStop'' (\citet{gsell2013sequential}). When applied to the max-t pvalues at FDR 0.10, {\tt forwardStop}
suggested stopping at step 8. In constant, the saturated lvalues gave stopping a point of 3, with no guarantee of FDR control
due to the lack of independence of the p-values.

An outline of this paper is as follows:
In Section \ref{sec:genericSetting} we setup the problem and give some important definitions. In Section \ref{sec:sequential} we review recently proposed methods for
testing ordered hypothesis using from a sequence of p-values. These method require that the p-values be uniform and independent under the null, and we derive general conditions
to ensure this property. In Section \ref{sec:singleStep}  we  show how to construct valid selective $p$-values for a single step in the path and derive the max-t test
for  forward stepwise regression and the Next-$\lambda$ test for the lasso. Section \ref{sec:computation}  describes our computational strategy for estimating the p-values using accept-reject and hit and run sampling. We also relate the method to parametric bootstrap sampling. We provide a simulation study in Section \ref{sec:sparseReg} and briefly discuss applications of our framework to
principal components analysis in section\ref{sec:pca}. The paper ends with a Discussion.

@@

\medskip
\begin{example}[The selective max-Z test]
Consider the forward stepwise path (\ref{eq:forwardDef_start}---\ref{eq:forwardDef_end}) in the case of linear regression, that is 
$\ell(\theta,Y)=||Y-X\theta||^2$.  Assume that the error variance $\sigma^2$ is known.
 Given an active set of variables $E$ in our current model $M$, the partial regression coefficients for predictors not in $E$ are
\begin{equation}
  z_{j,E}(Y) = \frac{(\proj_E^\perp X_j)'Y}
  {\sigma\|\proj_{E}^\perp X_j\|_2}.
\end{equation}
The max-$z$-statistic for active set $E$  is
\begin{equation}\label{eq:maxT}
  U_E(Y) = \max_{j \notin E} \left|t_{j,E}(Y)\right|,
\end{equation}
and we use $U_{E_{k-1}}(Y)$ as our test statistic at step $k$. We compute a p-value for this test
using the conditional law 
\begin{equation}\label{eq:selModel_marg0}
\L\left(X_{j_k}'Y \mid E_{k-1}, \; X_{E_{k-1}}'Y\right),
\end{equation}
If we reject the hypothesis, we add to $E$ the predictor that maximized $U_E(Y)$ and continue.
This is the procedure that used to obtain the p-values in Table \ref{tab:diab}, except that we sampled
conditioning on $||Y||$ as well (since $\sigma^2$ is unknown).
